The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-12-02 script/de/chapters/paper_and_pencil.tex
Script[de+en]: small corrections like index consistency.

%------------------------------------------------------------------------------
\section{Substitutionsverfahren}
\label{PaP_substitution_ciphers}
\index{Substitution}

@@

\newpage % Damit die Tables immer direkt nach dem zugehörigen Item. 
\item {\bf Codierung} \cite{pp:Singh2001}: Im Laufe der Geschichte wurden immer
   wieder Codebücher verwendet. In diesen Büchern wird jedem möglichen 
   {\bf Wort} eines Klartextes ein Codewort, ein Symbol oder eine Zahl
   zugeordnet.
   Voraussetzung für eine erfolgreiche geheime Kommunikation ist, dass 
   Sender und Empfänger exakt das gleiche Codebuch besitzen und die 
   Zuordnung der Codewörter zu den Klartextwörtern nicht offengelegt wird.

@@

%------------------------------------------------------------------------------
\section{Kombination aus Substitution und Transposition}
\label{PaP_combined_ciphers}
\index{Kaskaden}

@@

Die ursprüngliche Caesar-Chiffre ist einfach eine Shift-Chiffre,
wo der Schlüssel genau 3 ist. Im nächsten Beispiel nutzen wir die
Shift-Chiffre, um die Caesar-Chiffre über den Großbuchstaben
des lateinischen Alphabets zu bauen. % Dann verschlüsseln wir den
% Klartext P mit der Caesar-Chiffre. Schließlich entschlüsseln wir
% wieder das Chiffrat C und überprüfen, dass das Egebnis (DC)
% identisch mit dem ursprünglichen Klartext ist.

@@

\bibitem[Nguyen2009]{pp:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009, \\
       Siehe Sage publications page
       \url{http://www.sagemath.org/library-publications.html}

>>>>

%------------------------------------------------------------------------------
% \newpage
\clearpage 
\section{Substitutionsverfahren}
\label{PaP_substitution_ciphers}
\index{Substitution}

@@

\item {\bf Codierung} \cite{pp:Singh2001}: Im Laufe der Geschichte wurden immer
   wieder Codebücher verwendet. In diesen Büchern wird jedem möglichen 
   {\bf Wort} eines Klartextes ein Codewort, ein Symbol oder eine Zahl
   zugeordnet.
   Voraussetzung für eine erfolgreiche geheime Kommunikation ist, dass 
   Sender und Empfänger exakt das gleiche Codebuch besitzen und die 
   Zuordnung der Codewörter zu den Klartextwörtern nicht offengelegt wird.

@@

%------------------------------------------------------------------------------
\section{Kombination aus Substitution und Transposition}
\label{PaP_combined_ciphers}
\index{Kaskaden}
\index{Verschlüsselung!Mehrfachverschlüsselung}\index{Mehrfachverschlüsselung}

@@

Die ursprüngliche Caesar-Chiffre ist eine Shift-Chiffre mit
festem Schlüssel 3. Im nächsten Beispiel wird die Caesar-Chiffre
über den Großbuchstaben des lateinischen Alphabets gebaut.
% Dann verschlüsseln wir den
% Klartext P mit der Caesar-Chiffre. Schließlich entschlüsseln wir
% wieder das Chiffrat C und überprüfen, dass das Egebnis (DC)
% identisch mit dem ursprünglichen Klartext ist.

@@

\bibitem[Nguyen2009]{pp:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009, \\
       Siehe Sage publications
       \url{http://www.sagemath.org/library-publications.html}

### 2009-12-02 script/de/chapters/what-is-sage.tex
Script[de+en]: small corrections like index consistency.

Wenn man eigene Funktionen entwickelt, sie ändert und aufruft, dann ist es viel einfacher,
die Entwicklung in einem eigenen Editor vorzunehmen, den Code als Sage-Skriptdatei zu
speichern und die Funktionen nicht-interaktiv auf der Kommandozeile auszuführen.
Beide Arten, Code zu entwickeln, wurde in den Beispielen in
Kapitel \ref{PaP_Sage_samples} (\glqq \nameref{PaP_Sage_samples}\grqq) und in
Kapitel \ref{NumberTheory_Appendix_E} (\glqq\nameref{NumberTheory_Appendix_E}\grqq) angewandt.

@@

% \vspace{30pt}
\newpage
\noindent Weitere Hinweise:
\begin{itemize}
  \item Abfrage der Version Ihrer Sage-Umgebung mit: \texttt{version()}
  \item Um sich schnell die Sage-Programmbeispiele in diesem Skript anzusehen, können Sie
    \begin{itemize}
      \item im Index nach \verb#Sage -> Programmbeispiele# schauen, oder 
      \item sich im Anhang das \glqq \nameref{sc:List-of-Sage-Code-Examples}\grqq~ansehen.
    \end{itemize}
  \item Der Quellcode der Sage-Beispiele in diesem Script wird in Form von
        Sage-Programmdateien mit dem CrypTool-Setup-Programm ausgeliefert.
        Nach der Installation finden Sie diese im Unterverzeichnis \verb#sage# innerhalb des
        CrypTool-Verzeichnisses:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}

>>>>

Wenn man eigene Funktionen entwickelt, sie ändert und aufruft, dann ist es viel einfacher,
die Entwicklung in einem eigenen Editor vorzunehmen, den Code als Sage-Skriptdatei zu
speichern und die Funktionen nicht-interaktiv auf der Kommandozeile auszuführen.
Beide Arten, Code zu entwickeln, wurde in den Beispielen in
Kapitel \ref{PaP_Sage_samples} (\glqq \nameref{PaP_Sage_samples}\grqq),
Kapitel \ref{Primes_Sage_samples} (\glqq \nameref{Primes_Sage_samples}\grqq)
und in Kapitel \ref{NumberTheory_Appendix_E} (\glqq\nameref{NumberTheory_Appendix_E}\grqq)
angewandt.

@@

% \vspace{30pt}
\newpage
\noindent Weitere Hinweise:
\begin{itemize}
  \item Abfrage der Version Ihrer Sage-Umgebung mit: \texttt{version()}
  \item Um sich schnell die Sage-Programmbeispiele in diesem Skript anzusehen, können Sie
    \begin{itemize}
      \item im Index nach \verb#Sage -> Programmbeispiele# schauen, oder 
      \item sich im Anhang das \glqq \nameref{sc:List-of-Sage-Code-Examples}\grqq~ansehen.
    \end{itemize}
  \item Der Quellcode der Sage-Beispiele in diesem Skript wird in Form von
        Sage-Programmdatei"-en mit dem CrypTool-Setup-Programm ausgeliefert.
        Nach der Installation von CrypTool 1.x \index{CrypTool 1.x} finden Sie
        diese im Unterverzeichnis \verb#sage# innerhalb des
        CrypTool-Verzeichnisses:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}

### 2009-12-02 script/de/script-de.tex
Script[de+en]: small corrections like index consistency.

\vspace{30pt}
\begin{itemize}
  \item Alle Beispiel wurden mit Sage v 4.2.1 (Release Date 2009-11-14) getestet.
  \item Der Quellcode der Sage-Beispiele in diesem Script wird in Form von
        Sage-Programmdateien mit dem CrypTool-Setup-Programm ausgeliefert.
        Nach der Installation finden Sie diese im Unterverzeichnis \verb#sage# innerhalb des
        CrypTool-Verzeichnisses:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}

>>>>

\vspace{30pt}
\begin{itemize}
  \item Alle Beispiel wurden mit Sage v 4.2.1 (Release Date 2009-11-14) getestet.
  \item Der Quellcode der Sage-Beispiele in diesem Skript wird in Form von
        Sage-Programmdatei"-en mit dem CrypTool-Setup-Programm ausgeliefert.
        Nach der Installation von CrypTool 1.x \index{CrypTool 1.x} finden Sie
        diese im Unterverzeichnis \verb#sage# innerhalb des
        CrypTool-Verzeichnisses:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}

### 2009-12-02 script/en/chapters/authors.tex
Script[de+en]: small corrections like index consistency.

\subsection{For Grownups}

@@

\vskip +20 pt
Further samples of cryptology in fictional literature can be found on the
following German web page:
\begin{center}
    \href{http://www.staff.uni-mainz.de/pommeren/Kryptologie99/Klassisch/1\_Monoalph/Literat.html}
   {\texttt{http://www.staff.uni-mainz.de/pommeren/Kryptologie99/Klassisch/1\_Monoalph/}}\\
    \href{http://www.staff.uni-mainz.de/pommeren/Kryptologie99/Klassisch/1\_Monoalph/Literat.html}{\texttt{Literat.html}}
\end{center}
For some older authors (e.g. Jules Verne, Karl May, Arthur Conan Doyle,
Edgar Allen Poe) there are links to the original and relevant text pieces.\\

\newpage
\subsection{For Kids}

@@

\end{description}
You can find title pages of many of these books on the web site of Tobias Schr\"odel, who collects classic books about cryptography:\\
\href{http://tobiasschroedel.com/crypto_books.php}{\texttt{http://tobiasschroedel.com/crypto\_books.php}}

If you know of further literature and movies, where cryptography has a major
role or if you know of further books, which address cryptography in a
didactic and for children adequate way, then we would be very glad if
you could send us the exact book title and a short explanation
about the book's content. Thanks a lot.

>>>>

% --------------------------------------------------------------------------
\subsection{For Grownups}

@@

\vskip +20 pt

\paragraph{Remark 1:}
Further samples of cryptology in fictional literature can be found on the
following German web page:
\begin{center}
    \url{http://www.staff.uni-mainz.de/pommeren/Kryptologie99/Klassisch/1\_Monoalph/Literat.html}
\end{center}
For some older authors (e.g. Jules Verne, Karl May, Arthur Conan Doyle,
Edgar Allen Poe) there are links to the original and relevant text pieces.\\

\paragraph{Remark 2:}
You can find title pages of some of these books on the web site of
Tobias Schr\"odel, who collects classic books about cryptography:\\
\url{http://tobiasschroedel.com/crypto_books.php}

\paragraph{Remark 3:}
If you know of further books and movies, where cryptography has a major
role then we would be very glad if you could send us the exact title and
a short explanation about the movie/book's content. Thanks a lot.

% --------------------------------------------------------------------------
\newpage
\subsection{For Kids}

@@

\paragraph{Remark 1:}
You can find title pages of many of these kid books on the web site of
Tobias Schr\"odel, who collects classic books about cryptography:\\
\url{http://tobiasschroedel.com/crypto_books.php}

\paragraph{Remark 2:}
If you know of further books, which address cryptography in a
didactic and for children adequate way, then we would be very glad if
you could send us the exact book title and a short explanation
about the book's content. Thanks a lot.

### 2009-12-02 script/en/chapters/cryptomethods.tex
Script[de+en]: small corrections like index consistency.

More details about the AES algorithms and the AES candidates of the last round
can be found within the online help of CrypTool\index{CrypTool}%
\footnote{%
      CrypTool online help\index{CrypTool}: the index head-word {\bf AES}
      leads to the 3 help pages: {\bf AES candidates}, 
      {\bf The AES winner Rijndael} and 
      {\bf The Rijndael encryption algorithm}.
  }.

@@

The mathematicians Courtois and Pieprzyk \cite{cm:Courtois2002} published a paper
in 2002, which got a great deal of attention amongst the cryptology community: The
pair had further developed the XL-method (eXtended Linearization), introduced at
Eurocrypt 2000 by Shamir et al., to create the so called XSL-method (eXtended
Sparse Linearization). The XL-method is a heuristic technique, which in some
cases manages to solve big non-linear equation systems and which was till then
used to analyze an asymmetric algorithm (HFE).  The innovation of Courtois and
Pieprzyk was, to apply the XL-method on symmetric codes: the XSL-method can be
applied to very specific equation systems. A 256-bit AES could be attacked in
roughly $2^{230}$ steps. This is still a purely academic attack, but also a
direction pointer for a complete class of block ciphers. The major problem with
this attack is that until now nobody has worked out, under what conditions it is
successful: the authors specify in their paper necessary conditions, but it is
not known, which conditions are sufficient.  There are two very new aspects of
this attack: firstly this attack is not based on statistics but on algebra. So
attacks seem to be possible, where only very small amounts of ciphertext are
available. Secondly the security of a product algorithm\index{Cascade cipher}%
\footnote{%
A ciphertext can be used as input for another encryption algorithm. 
A cascade cipher is build up as a composition of different encryption 
transformations. The overall cipher is called product algorithm or cascade cipher
(sometimes depending whether the used keys are statistically dependent or not).\\
Cascading does not always improve the security.\\
This process is also used {\em within} modern algorithms:
They usually combine simple and, considered at its own, cryptologically
relatively insecure single steps in several rounds into an efficient 
overall procedure.  Most block ciphers (e.g. DES, IDEA) are cascade ciphers.\\
Also serial usage of the same cipher with different keys (like with Triple-DES)
is called cascade cipher. 
}
does not exponentially increase with the number of rounds.

@@

Companies like RSA Security provide so-called cipher challenges in order to quantify the security offered by well-known symmetric ciphers as DES, Triple-DES or RC5\footnote{%
 \url{http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}\\
 In May 2007 RSA announced that they will not confirm the correctness of
 the not yet solved RC5-72 challenge.}.
They offer prizes for those who manage to decipher ciphertexts, encrypted with different algorithms and different key lengths, and to unveil the symmetric key (under controlled conditions). So theoretical estimates can be confirmed.

@@

One of these variants is e.g. S-AES (Simplified-AES) by Prof. Ed Schaefer
and his students \cite{cm:Musa-etal2003}%
\footnote{
    See the articles
    ``Devising a Better Way to Teach and Learn the Advanced Encryption Standard''\\
    at the university's news at\\
    \url{http://math.scu.edu/~eschaefe/getfile.pdf} \\
    \url{http://www.scu.edu/cas/research/cryptography.cfm}
}.
Another one is Mini-AES \cite{cm:Phan2002}
(see chapter~\ref{CM_Sage_Mini-AES} ``\nameref{CM_Sage_Mini-AES}''):
\begin{itemize}

@@

@@

\bibitem[Nguyen2009]{cm:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009, \\
       See Sage publications page
       \url{http://www.sagemath.org/library-publications.html}

>>>>

More details about the AES algorithms and the AES candidates of the last round
can be found within the online help of CrypTool\index{CrypTool}%
\footnote{%
      CrypTool online help\index{CrypTool}: The index head-word {\bf AES}
      leads to the 3 help pages: {\bf AES candidates}, 
      {\bf The AES winner Rijndael} and 
      {\bf The Rijndael encryption algorithm}.
  }.

@@

The mathematicians Courtois and Pieprzyk \cite{cm:Courtois2002} published a paper
in 2002, which got a great deal of attention amongst the cryptology community: The
pair had further developed the XL-method (eXtended Linearization), introduced at
Eurocrypt 2000 by Shamir et al., to create the so called XSL-method (eXtended
Sparse Linearization). The XL-method is a heuristic technique, which in some
cases manages to solve big non-linear equation systems and which was till then
used to analyze an asymmetric algorithm (HFE).  The innovation of Courtois and
Pieprzyk was, to apply the XL-method on symmetric codes: the XSL-method can be
applied to very specific equation systems. A 256-bit AES could be attacked in
roughly $2^{230}$ steps. This is still a purely academic attack, but also a
direction pointer for a complete class of block ciphers. The major problem with
this attack is that until now nobody has worked out, under what conditions it is
successful: the authors specify in their paper necessary conditions, but it is
not known, which conditions are sufficient.  There are two very new aspects of
this attack: firstly this attack is not based on statistics but on algebra. So
attacks seem to be possible, where only very small amounts of ciphertext are
available. Secondly the security of a product algorithm%
\index{Product algorithm}\index{Encryption!Product algorithm}%
\index{Cascade cipher}\index{Encryption!Cascade cipher}%
\footnote{%
A ciphertext can be used as input for another encryption algorithm. 
A cascade cipher%
\index{Product algorithm}\index{Encryption!Product algorithm}%
\index{Cascade cipher}\index{Encryption!Cascade cipher}%
is build up as a composition of different encryption transformations.
The overall cipher is called product algorithm or cascade cipher
(sometimes depending whether the used keys are statistically dependent or not).\\
Cascading does not always improve the security.\\
This process is also used {\em within} modern algorithms:
They usually combine simple and, considered at its own, cryptologically
relatively insecure single steps in several rounds into an efficient 
overall procedure.  Most block ciphers (e.g. DES, IDEA) are cascade ciphers.\\
Also serial usage of the same cipher with different keys (like with Triple-DES)
is called cascade cipher. 
}
does not exponentially increase with the number of rounds.

@@

Companies like RSA Security provide so-called cipher challenges in order
to quantify the security offered by well-known symmetric ciphers as DES,
Triple-DES or RC5%
\index{DES}\index{RC5}\index{Triple-DES}%
\footnote{%
 \url{http://www.rsasecurity.com/rsalabs/challenges/secretkey/index.html}\\
 In May 2007 RSA announced that they will not confirm the correctness of
 the not yet solved RC5-72 challenge.}.
They offer prizes for those who manage to decipher ciphertexts, encrypted with different algorithms and different key lengths, and to unveil the symmetric key (under controlled conditions). So theoretical estimates can be confirmed.

@@

One of these variants is e.g. S-AES (Simplified-AES) by Prof. Ed Schaefer
and his students \cite{cm:Musa-etal2003}%
\footnote{
    See the two articles with the same title
    ``Devising a Better Way to Teach and Learn the Advanced Encryption Standard''
    at the university's news at\\
    - \url{http://math.scu.edu/~eschaefe/getfile.pdf} \\
    - \url{http://www.scu.edu/cas/research/cryptography.cfm}
}.
Another one is Mini-AES \cite{cm:Phan2002}
(see chapter~\ref{CM_Sage_Mini-AES} ``\nameref{CM_Sage_Mini-AES}'')%
\index{DES}\index{DES!SDES}\index{IDEA}:
\begin{itemize}

@@

Further example code for Mini-AES can be found in \cite[chap. 6.5 and appendix D]{cm:Nguyen2009}.

@@

\bibitem[Nguyen2009]{cm:Nguyen2009} \index{Nguyen 2009} 
       Minh Van Nguyen, \\
       {\em Exploring Cryptography Using the Sage Computer Algebra System}, \\
       Victoria University, 2009, \\
       See Sage publications
       \url{http://www.sagemath.org/library-publications.html}

### 2009-12-02 script/en/chapters/digitalsignatures.tex
CT script: Minor corrections, mostly layout, de + en; final now for CT 1.4.30 Beta04
Script[de+en]: small corrections like index consistency.

Cryptographically secure hash functions fulfill the following requirements
(the order is in a way that the requirements increase):
\begin{itemize}
 \item Resistance against 1st Pre-Image attacks:
      \index{Pre-Image-Attack!1st}  \\
      It should be practically impossible, for a given number, to find a
      message that has precisely this number as hash value. \\
      Given (fix): hash value H', \\
      Searched: message m, so that: H(m) = H'.
 \item Resistance against 2nd Pre-Image attacks:
      \index{Pre-Image-Attack!2nd}  \\
      It should be practically impossible, for a given message, to find another
      message, which has precisely the same hash value. \\
      Given (fix): message m1 [and so the hash value H1 = H(m1)], \\
      Searched: message m2, so that: H(m2) = H1.
 \item Collision resistance:
      \index{Collision resistance}  \\
      It should be practically impossible to find any two messages with the
      same hash value (it doesn't matter what hash value). \\
      Searched: 2 messages m1 and m2, so that: H(m1) = H(m2).
\end{itemize}

@@

The initially published result reduced the expected complexity for one SHA-1
collision search from $2^{80}$ (brute force) to $2^{69}$
\cite{ds:Wang2005}. More recent announcements claim to further reduce the required
effort to $2^{63}$ \cite{ds:Wang2005b} and $2^{52}$ \cite{ds:McDonald2009}. This would
bring collision attacks into the practical realm, as similar efforts have been
mastered in the past (s.\ \ref{Brute-force-gegen-Symmetr}).

>>>>

Cryptographically secure hash functions fulfill the following three requirements
(the order is in a way that the requirements increase):
\begin{itemize}
 \item Resistance against 1st Pre-Image attacks:
      \index{Pre-Image-Attack!1st}  \\
      It should be practically impossible, for a given number, to find a
      message that has precisely this number as hash value. \\
      Given (fix): hash value H', \\
      Searched: message m, so that: H(m) = H'.
 \item Resistance against 2nd Pre-Image attacks:
      \index{Pre-Image-Attack!2nd}  \\
      It should be practically impossible, for a given message, to find another
      message, which has precisely the same hash value. \\
      Given (fix): message m1 [and so the hash value H1 = H(m1)], \\
      Searched: message m2, so that: H(m2) = H1.
 \item Collision resistance:
      \index{Collision resistance}  \\
      It should be practically impossible to find any two messages with the
      same hash value (it doesn't matter what hash value). \\
      Searched: 2 messages m1 and m2, so that: H(m1) = H(m2).
\end{itemize}

@@

The initially published result reduced the expected complexity for one SHA-1
collision search from $2^{80}$ (brute-force) to $2^{69}$
\cite{ds:Wang2005}. More recent announcements claim to further reduce the required
effort to $2^{63}$ \cite{ds:Wang2005b} and $2^{52}$ \cite{ds:McDonald2009}. This would
bring collision attacks into the practical realm, as similar efforts have been
mastered in the past (s.\ \ref{Brute-force-gegen-Symmetr}).

### 2009-12-02 script/en/chapters/ellipticcurves.tex
Script[de+en]: small corrections like index consistency.

(Note that $-(-P)=(x,x+(x+y))=(x,2x+y)=(x,y)$, since the underlying field has characteristic $2$.)\footnote{An animation of the addition of points on elliptic curves can be found on the Certicom\index{Certicom}-Homepage \\
\href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\texttt{http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}}}

@@

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  According to the Sage-support newsgroup there is work underway
  to create a system for using @interact completely outside of the
  Sage notebook (JS code within a static html pages).

@@

% -----------------------------------------------------------------------------
\newpage
\section{Patent aspects}

@@

\begin{enumerate}
        \hypertarget{ec:Web-Link:Sage_Massierer}{} 
   \item Interactive introduction to elliptic curves and elliptic curve cryptography
        with Sage\index{Sage} by Maike Massierer and the CrypTool team,\\
        \href{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/}
        {\texttt{http://sage.mathematik.uni-siegen.de:8000/home/pub/45/}} \\
        ECC Tutorial as Sage Notebook \\
        Version 1.01, October 2008
   \item Certicom Online Tutorial, \index{Certicom},\\
        \href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}
        {\texttt{http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}}

   \item Working group IEEE P1363, \\
        \href{http://grouper.ieee.org/groups/1363}
        {\texttt{http://grouper.ieee.org/groups/1363}}

   \item \hypertarget{Lenstra2}{}
        An informative web page about factorization with elliptic curves, \\
        \href{http://www.loria.fr/~zimmerma/records/ecmnet.html}
	{\texttt{http://www.loria.fr/\~{}zimmerma/records/ecmnet.html}} \\
        It contains literature related to the topic factorization with 
	elliptic curves as well as links to other web page.

   \item Key length comparison by Arjen Lenstra and Eric Verheul,\\
        \href{http://cryptosavvy.com/table.htm}
        {\texttt{http://cryptosavvy.com/table.htm}}

>>>>

(Note that $-(-P)=(x,x+(x+y))=(x,2x+y)=(x,y)$, since the underlying field
has characteristic $2$.)\footnote{An animation of the addition of points
on elliptic curves can be found on the Certicom\index{Certicom} homepage \\
\url{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}}

@@

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  According to the Sage-support newsgroup there is work underway
  to create a system for using \verb#@interact# completely outside of the
  Sage notebook (JS code within a static html pages).

@@

% -----------------------------------------------------------------------------
\newpage
\section{Patent aspects}
\index{Patent}

@@

\begin{enumerate}

        \hypertarget{ec:Web-Link:Sage_Massierer}{} 
   \item Interactive introduction to elliptic curves and elliptic curve cryptography
        with Sage\index{Sage} by Maike Massierer and the CrypTool team,\\
        \url{http://sagenb.org/home/pub/1126/} (\verb!#1126 to #1133!) \\
        ECC Tutorial as Sage Notebook \\
        Version 1.2, November 2009

   \item Certicom Online Tutorial\index{Certicom},\\
        \url{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}

   \item Working group IEEE P1363, \\
        \url{http://grouper.ieee.org/groups/1363}

   \item \hypertarget{Lenstra2}{}
        An informative web page about factorization with elliptic curves, \\
        \url{http://www.loria.fr/~zimmerma/records/ecmnet.html}\\
        It contains literature related to the topic factorization with 
	elliptic curves as well as links to other web page.

   \item Key length comparison by Arjen Lenstra and Eric Verheul,\\
        \url{http://cryptosavvy.com/table.htm}

### 2009-12-02 script/en/chapters/introduction.tex
CT script: Sage parts finished in (en); translation to German almost finished
Script[de+en]: Many small corrections -- should be close to final now.
Script[de+en]: small corrections like index consistency.

The rapid spread of the Internet has also lead to intensified research in the
technologies involved, especially within the area of cryptography where a good
deal of new knowledge has arisen.

%This edition of the script adds some topics, but mainly updates areas (e.g. the
%summaries of topical research areas):
This edition completely updated the TeX sources of the document, and of course
the content of the script was corrected, amended and updated with some topics, e.g.:
\vspace{-7pt}
\begin{itemize}
  \item the search for the largest prime numbers 
        (chap. \ref{search_for_very_big_primes}), 
  \item progress in cryptanalysis of hash algorithms 
        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played major role (see appendix \ref{s:appendix-movies});
        and where primes are used as hangers  
        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}).
\end{itemize}

Newly added is the appendix \ref{s:appendix-using-sage} about using the
computer algebra system Sage. This was added because Sage becomes
more and more the standard open-source CAS system. Accordingly all
samples written before in Pari-GP and Mathematica have been substituted
with Sage code. Thanks to Minh Van Nguyen, a lot of new code samples could be added.

@@

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\\
\\
% [1.5\baselineskip]
% \enlargethispage*{2\baselineskip}
% \nopagebreak
Bernhard Esslinger
\\
\\
% [\baselineskip]
Frankfurt (Germany), August 2009

@@

At this point I'd like to thank explicitly the following people who
particularly contributed to CrypTool\index{CrypTool}. They applied their
very special talents and showed really great engagement:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Staff of Prof.\ Claudia Eckert, Prof.\ Johannes Buchmann and Prof.\ Torben Weis.
\end{list}
%\end{itemize}
Also I want to thank all the many people not mentioned here for their 
hard work (mostly carried out in their spare time).
\\
\\
Bernhard Esslinger
\\
\\
Frankfurt (Germany), July 2009

>>>>

The rapid spread of the Internet has lead to intensified research in the
technologies involved, especially within the area of cryptography where a good
deal of new knowledge has arisen.

%This edition of the script adds some topics, but mainly updates areas (e.g. the
%summaries of topical research areas):
This edition completely updated the TeX sources of the document, and of course
the content of the script was corrected, amended and updated with some topics, e.g.:
\vspace{-5pt}
\begin{itemize}
  \item the search for the largest prime numbers 
        (chap. \ref{search_for_very_big_primes}), 
  \item progress in cryptanalysis of hash algorithms 
        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played major role (see appendix \ref{s:appendix-movies});
        and where primes are used as hangers  
        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}).
  \item Newly added is the appendix \ref{s:appendix-using-sage} about using the
        computer algebra system Sage. This was added because Sage becomes
        more and more the standard open-source CAS system. Accordingly all
        samples written before in Pari-GP and Mathematica have been substituted
        with Sage code. Thanks to Nguyen and Massierer, a lot of new code samples
        could be added (see also chap.~\ref{ec:Sage_Massierer}).
\end{itemize}

@@

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\\
\\
% [1.5\baselineskip]
% \enlargethispage*{2\baselineskip}
% \nopagebreak
Bernhard Esslinger
\\
\\
% [\baselineskip]
Frankfurt (Germany), December 2009

@@

At this point I'd like to thank explicitly the following people who
particularly contributed to CrypTool\index{CrypTool}. They applied their
very special talents and showed really great engagement:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Staff of Prof.\ Claudia Eckert, Prof.\ Johannes Buchmann and Prof.\ Torben Weis.
\end{list}
%\end{itemize}
Also I want to thank all the many people not mentioned here for their 
hard work (mostly carried out in their spare time).
\\
\\
Bernhard Esslinger
\\
\\
Frankfurt (Germany), November 2009

### 2009-12-02 script/en/chapters/menus.tex
1.4.30 menu tree update
Script[de+en]: small corrections like index consistency.

This appendix contains at the following page the complete menu tree of
CrypTool\index{CrypTool} version 1.4.30\footnote{%
  The future versions CrypTool 2 and JCrypTool of the CrypTool project are
  currently there only as (stable) beta versions. If the according release
  versions are available, we will add the appropriate menus and menu trees.\\
}.

Which menu items in CrypTool are active (that means not greyed), depends on
the type of the currently active document window.

The brute-force analysis\index{Attack!brute-force} for DES e.~g. is only
available, if the active window is opened in the hexadecimal view. 
On the other hand the menu item ``Generate Random Numbers\dots''
is always available (even if no document is opened).

@@

%\nobreak
\clearpage
\begin{figure}[hb]
\begin{center}
\vspace{-30pt}
%\frame{  %TeX macht einen Rahmen (siehe viewport) drumherum -- gut zum Testen
%\includegraphics[scale=0.25, angle=270, viewport=200 30 2660 1430]
\includegraphics[scale=0.25, angle=270]
                {figures/cryptool-menu-en}
%viewport=rand-links? rand-unten breite hoehe? [bezogen auf querformat]
%}
\caption{Complete overview of the menu tree of CrypTool 1.4.30} 
\label{menuoverview}
\end{center}
\end{figure}
\clearpage

>>>>

This appendix contains at the following page the complete menu tree of
CrypTool\index{CrypTool} version 1.4.30\footnote{%
  In parallel to CrypTool 1.x\index{CrypTool 1.x} the future versions
  CrypTool 2\index{CrypTool 2.0} and JCrypTool\index{JCrypTool 1.0}
  are currently developed in the CrypTool project.\\
  - Webseite CT2: \url{http://www.cryptool2.vs.uni-due.de} \\
  - Webseite JCT: \url{http://jcryptool.sourceforge.net} \\
  These future versions are currently (December 2009) beta versions, but they
  are stable enough to be used by end-users already. If the according release
  versions are available, we will add the appropriate menus and menu trees.\\
}.

\noindent The main menu contains service functions in the menus
\begin{itemize}
   \item File
   \item Edit
   \item View
   \item Options
   \item Window
   \item Help
\end{itemize}
and the actual cyrpto functions in the menus
\begin{itemize}
   \item Crypt/Decrypt
   \item Digital Signature/PKI
   \item Individual Procedures
   \item Analysis.
\end{itemize}

Below \verb#Individual Procedures# you find visualizations of single algorithms
and of protocols. Some procedures are implemented both for a fast performance
(mostly under the main menu \verb#Crypt/Decrypt#) and for a step-by-step visualization.

Which menu items in CrypTool are active (that means not greyed), depends on
the type of the currently active document window:
The brute-force analysis\index{Attack!brute-force} for DES e.~g. is only
available, if the active window is opened in the hexadecimal view. 
On the other hand the menu item ``Generate Random Numbers\dots''
is always available (even if no document is opened).

@@

%\nobreak
\clearpage
\begin{figure}[hb]
\begin{center}
\vspace{-30pt}
%\frame{  %TeX macht einen Rahmen (siehe viewport) drumherum -- gut zum Testen
%\includegraphics[scale=0.25, angle=270, viewport=200 30 2660 1430]
\includegraphics[scale=0.35, angle=270]
                {figures/cryptool-menu-en}
%viewport=rand-links? rand-unten breite hoehe? [bezogen auf querformat]
%}
\caption{Complete overview of the menu tree of CrypTool 1.4.30} 
\label{menuoverview}
\end{center}
\end{figure}
\clearpage
